Off-target spray particle movement

### 2021-08-04 manuscript.Rmd
as

---
title: "Off-target spray particle movement"
author: 
  - "Rodrigo Werle"
  - "Bruno C Vieira"
  - "Maxwel C. Oliveira"
  - "Guilherme Alves"
  - "Greg Kruger"
date: "`r format(Sys.Date())`"
documentclass: article
linestretch: 2
abstract: "BACKGROUND: "
papersize: "a4"
mainfont: Arial
#monofont: "Arial"
#monofontoptions: "Scale=0.7"
fontsize: "12pt"
geometry:
  margin = 2.54cm
csl: pms.csl
bibliography: biblio.bib
output:
  pdf_document: 
      latex_engine: xelatex
      keep_tex: true
      toc_depth: 4
      number_sections: true

@@

Integrate pest management The off-target particle movement

The introduction of

The objective of this study was to compare

@@

@@

>>>>

---
title: "Off-target spray particle movement"
author: 
  - "Rodrigo Werle"
  - "Bruno C Vieira"
  - "Maxwel C. Oliveira"
  - "Guilherme Alves"
  - "Greg Kruger"
date: "`r format(Sys.Date())`"
documentclass: article
linestretch: 2
abstract: "BACKGROUND: "
papersize: "a4"
mainfont: Arial
#monofont: "Arial"
#monofontoptions: "Scale=0.7"
fontsize: "12pt"
geometry:
  margin = 2.54cm
csl: pms.csl
bibliography: spray-pms.bib
output:
  pdf_document: 
      latex_engine: xelatex
      keep_tex: true
      toc_depth: 4
      number_sections: true

@@

The widespread cases of glyphosate-resistant weeds throughout the US motivated farmers to adopt dicamba and 2,4-D tolerant crops [@behrens2007; @heap2018; @sammons2014; @werle2018; @wright2010]. As expected, the agricultural usage of dicamba and 2,4-D significantly increased in the US (“United States Department of Agriculture” 2020). Applications of dicamba and 2,4-D in new cropping areas during extended time in the growing season has led to increased opportunities for the off-target movement of these herbicides [@soltaniOfftargetMovementAssessment2020; @werle2021]. Spray drift is one of the main mechanisms of pesticide off-target movement and can be defined as the part of the pesticide application deflected away from the target area during or following applications [@matthews2008]. Herbicide drift potential is influenced by application technique, environmental conditions, and surrounding vegetation [@alheidary2014; @arvidsson2011; @felsot2010; @hewitt2000; @nordby1974].

Dicamba and 2,4-D spray drift is a concern because these herbicides are very active on broadleaf vegetation, including soybean and cotton, causing distinctive crop injury and yield loss at very low dose exposures [@egan2014]. The adoption of best management practices during herbicide applications, including proper nozzle selection and the use of drift reducing adjuvants, are among the main strategies to manage spray droplet size and mitigate spray drift [@etheridge1999; @hewitt2008; @hilz2013]. New dicamba and 2,4-D products are mostly applied using venturi nozzles with air-inclusion and pre-orifice components that increase spray droplet size and reduce spray drift potential [@creech2015; @dorr2013]. Despite the advances in nozzle technology [@dorr2013], pesticide formulations and adjuvants [@hilz2013], spraying techniques [@butts2019], strategies to mitigate spray drift [@felsot2010], and education and extension efforts [@butts2021; @werle2018], dicamba and 2,4-D spray drift remains associated with crop injury complaints across the US. In addition to this, common practices to mitigate spray drift, such as the use of DRAs with venturi nozzles in combination with lower pressures and boom heights, could potentially compromise spray uniformity, deposition, and weed control [@wolf1993] (Moraes et al. 2021).

The use of spray shields or hoods is another strategy to mitigate spray drift during pesticide applications [@burnside1968; @foster2018; @ozkan1997; @d.b.smith1982; @wolf1993]. These devices generally reduce spray drift potential by minimizing spray exposure to wind. Foster et al. (2018)[@foster2018] reported that a hooded sprayer reduced spray drift potential in approximately 50% regardless of the nozzles used compared to conventional applications. The use of spray hoods can potentially extend the available time for applications during windy conditions and allow the use of finer droplets with lower carrier volumes to optimize spray coverage and weed control [@wolf1993]. Although the benefits of sprayer hoods have been studies and reported since the early fifties, the technique adoption has been relatively low among farmers and applicators [@nordby1974]. With the increased usage of dicamba and 2,4-D and the spray drift incidents associated with both herbicides, crop protection professionals renewed their interest in spray hoods to further mitigate herbicide off-target movement during applications. Therefore, the objective of this study was to evaluate the effectiveness of spray hoods in reducing pesticide drift for spray solutions and nozzles typically used for new auxinic herbicide products.

@@

## Droplet size

As expected, the TTI nozzle produced coarser spray classification compared to the ULD and the AIXR nozzles across all spray solutions tested. In a response surface modeling approach in which nozzles were tested at different operational setups, TTI nozzles had greater percentage of operational setups producing Ultra Coarse (90%) spray classification compared to ULD (2%) and AIXR (0%) nozzles [@fritz2016]. Legleiter et al. (2018)[@legleiter2018] reported that glyphosate + 2,4-D (premixture formulation) applications using a TTI11004 nozzle produced coarser spray classification (Ultra Coarse) when compared to AIXR11004 (Very Coarse), TT11004 (Coarse), and XR11004 (Medium) nozzles. Alves et al. (2020)[@alves2020] reported that applications of different dicamba formulations using a TTI110015 nozzle had larger droplet size across all formulations when compared to applications using an AIXR110015 nozzle. Zaric (2020)[@zaric2020] reported that the TTI11004 nozzle consistently produced coarser spray droplet size across multiple spray solutions when compared to the ULD12004 nozzle.

Creech et al. (2015)[@creech2015] reported that nozzle design had major effects on spray droplet size, whereas spray solution influenced droplet spectrum to a less extent. The addition of DRA increased spray droplet size across all nozzles tested herein. Samples et al. (2021)[@samples2021] reported that the addition of DRAs (polymer and guargum) increased the spray droplet size of herbicide spray mixtures containing combinations of glyphosate, glufosinate, acetochlor, *S*-metolachlor, dicamba, fomesafen, and/or 2,4-D applied with an AIXR11005 nozzle. Zaric (2020)[@zaric2020] reported that the addition of DRAs consistently increased spray droplet size for dicamba applications using TTI1104 and ULD12004 spray nozzles. These findings highlight the specific instructions regarding nozzle selection for the new dicamba products (“Engenia Product Label” 2021, “Xtendimax Product Label” 2021), in which nozzles producing Ultra Coarse spray classifications are approved for applications. Label instructions for new 2,4-D formulations are more flexible in terms of nozzle selection, where Coarse, Very Coarse, and Extremely Coarse spray classifications are allowed (“Enlist Duo Product Label” 2021, “Enlist One Product Label” 2021). New dicamba and 2,4-D product labels also include specific DRAs to be used, especially when products are tank mixed with other pesticides.

## Spray solution deposition in-swath

For conventional applications (open) without DRA, the TTI nozzle had greater in-swath spray deposition volume on petri dishes compared to the ULD and AIXR nozzles. It has been previously reported that applications with Ultra Coarse spray had greater in-swath deposition volume on petri dishes compared to finer spray classifications [@vieira2018]. Finer sprays have a greater percentage of spray volume comprised of small droplets that are more susceptible to spray drift [@dorr2013]. These small droplets have more aerodynamic drag and lose their initial momentum rapidly, therefore having their trajectory highly influenced by the wind compared to coarser droplets [@knoche1994]. It is important to highlight that greater spray deposition volume collected on petri dishes does not mean greater spray coverage, as coarser sprays result in reduced spray coverage when compared to finer spray classifications [@knoche1994; @legleiter2018].

For conventional applications (open), the addition of DRA consistently increased in-swath spray deposition volume (petri dishes) compared to applications without DRA. Samples et al. (2021) [@samples2021] reported that the addition of DRAs (polymer and guargum) to several herbicide tank mixtures had inconsistent deposition results on soybean and cotton when compared to herbicide applications without adjuvants. Authors reported that while the addition of polymer and guargam based adjuvants increased cotton spray deposition for glyphosate + dicamba + *S*-Metolachlor spray mixture, the same adjuvants had no effect or decreased cotton spray deposition when mixed to glyphosate + 2,4-D. Moraes et al. (2021) (Moraes et al. 2021) reported that the addition of DRAs could have a negative effect on spray uniformity, especially with nozzles producing coarser spray classifications at lower spray pressures.

Results reported herein indicate an inconsistent effect of hooded sprayers on in-swath spray deposition. For applications with DRA, the conventional sprayer (open) had greater in-swath deposition compared to hooded applications. On the other hand, for applications without DRA, the hooded sprayer had greater in-swath deposition compared to the conventional sprayer. Wolf et al. (1993) [@wolf1993] observed that solid spray shields generally decreased in-swath deposition uniformity. Authors noted that spray solution dripping from sprayer shield walls could increase the in-swath deposition variability during applications. Wolf et al (1993) [@wolf1993] and Moraes et al. (2021) observations could explain the lower spray deposition obtained for hooded applications with the addition of the DRA, although this hypothesis needs to be further investigated.

@@

For conventional (open) applications without DRA, the TTI nozzle had less drift potential compared to AIXR and ULD nozzles, as expected. Alves et al. (2020) [@alves2020] reported similar results in a wind tunnel study, where applications of different dicamba formulations using a TTI nozzle had less spray drift compared to applications using an AIXR nozzle. Authors reported that soybean plants were still showing 50% biomass reduction from 9.9 to 13.4 m downwind the nozzle for applications with the AIXR nozzle, whereas this distance range was reduced to 6.4 to 7.2 m for applications using the TTI nozzle. In another wind tunnel study, Alves et al. (2017) [@alves2017] reported that dicamba applications with a TTI nozzle had 0.6% spray drift at 12 m downwind the nozzle, whereas applications with XR, TT, and AIXR nozzles had 13.5, 6.6, and 3.2% drift, respectively. As previously reported [@hilz2013; @johnson2006], the addition of DRA reduced spray drift across all nozzles tested herein.

Hooded applications substantially reduced spray drift potential across all treatment scenarios. For instance, hooded applications using the AIXR nozzle without DRA had similar drift potential compared to conventional applications using the TTI nozzle with DRA despite the major droplet size differences between these treatments. These results indicate that the adoption of spray hoods combined with proper nozzle selection and the addition of DRAs to the spray solution can substantially reduce spray drift potential during pesticide applications. Foster et al. (2018)[@foster2018] reported that the advent of spray hood considerably reduced spray drift for glyphosate applications using Fine, Medium, Very Coarse, and Ultra Coarse spray qualities. In a wind tunnel study, Ozkan et al. (1997)[@ozkan1997] reported that several spray shields were effective in reducing spray drift, where a double-foil shield reduced spray drift in 59% compared to applications without shields. In laboratory tests, Smith et al. (1982) reported that a spray shield could reduce spray drift up to 70%, although field studies indicated an inconsistent performance where spray shields either increased or decreased spray drift during tests.

The results reported in this study highlight the importance of adopting best management practices during pesticide applications to mitigate spray off-target movement to the surrounding environment. Hooded applications with AIXR and ULD nozzles resulted in satisfactory spray drift reduction despite the finer droplet spectrum compared to the TTI nozzle. This could have implications for weed control as applications with Ultra Coarse droplet spectrum have reduced spray coverage compared to finer spray classifications. Legleiter et al. (2018)[@legleiter2018] reported that glyphosate + 2,4-D applications with Ultra Coarse (TTI11004) spray had reduced spray coverage on artificial targets (cards) compared to Medium (XR11004), Coarse (TT11004), and Very Coarse (AIXR11004) sprays. However, authors reported that spray droplet spectrum did not influence 2,4-D absorption in Palmer amaranth (*Amaranthus palmeri*), waterhemp (*Amaranthus tuberculatus*), giant ragweed (*Ambrosia trifida*), and horseweed (*Erigeron canadensis*). Increasing the application carrier volume is one strategy to overcome the reduced coverage from applications using coarser sprays. Butts et al. (2018) [@butts2018] reported that dicamba applications with Ultra Coarse spray resulted in optimal weed control for applications at 187 L ha-1, whereas Extremely Coarse spray resulted optimal weed control for glufosinate applications. However, applications with increased spray carrier volume are less efficient due to additional time and water requirements during the operation [@wolf1993]. Hooded applications using AIXR and ULD nozzles can provide increased spray coverage compared to TTI nozzles while having low drift potential, which allows for applications at reduced carrier volumes.

Because of the substantial drift reduction, applications with hooded sprayers could also reduce product buffer zone requirements. New dicamba formulations provide incentives to the use of qualified hooded sprayers by decreasing the required downwind buffer from 73 m to 33 m (“Engenia Product Label” 2021, “Xtendimax Product Label” 2021). Although the benefits associated with hooded sprayers, the technology has not been well accepted by farmers. Nordby and Skuterud (1974) [@nordby1974] mentioned that farmers could see spray hoods as cumbersome and expensive devices that do not allow applicators to check nozzles during applications. Considering the widespread adoption of dicamba and 2,4-D tolerant crops and the increase in these herbicide usage, hooded sprayers could be a very effective technique to mitigate spray drift during pesticide applications.
